Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2025-01-14 demystify.tex
Updates from Overleaf

\begin{mybox}
This book is open source and updated regularly. Its latest version is available at

@@

International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US. Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.

@@

\paragraph{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries. It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries). It doesn't even require having a CS background (\autoref{sec:non-stem}).
If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.

\paragraph{Funding is not an issue} Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. 
If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
health insurance, and stipend (i.e., in STEM field you get paid for your study!). 
For many students, the stipends provided are not only sufficient to cover your own living expenses but also enough to support your families (e.g., spouse and children), allowing them to live comfortably.

@@

\paragraph{Discrimination is not tolerated} Third, not only is discrimination illegal, diversity and inclusion are highly valued in US universities. Many universities take great pride in their diverse student body and faculty.

@@

This book also aims to explain the reasons behind the admission process. There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc. Understanding the reason and mindset of the admission committee and profs. can help you prepare better.

While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions, which are frankly the majority of good schools in the US.

>>>>

\begin{mybox}
This book is open source and available at

@@

International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US. Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, (iii) the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.

@@

\paragraph{Not any harder than other countries} Applying to a good US university \emph{should not} be any harder than at schools in other countries. It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries). It doesn't even require having a CS background (\autoref{sec:non-stem}).
If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.

\paragraph{Funding is not an issue} PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. 
If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
health insurance, and stipend (i.e., in STEM field you get paid for your study!). 
For many students, the stipends provided are not only sufficient to cover your own living expenses but also enough to support your families (e.g., spouse and children), allowing them to live comfortably.

@@

\paragraph{Discrimination is not tolerated} Not only is discrimination illegal, diversity and inclusion are highly valued in US universities. Many universities take great pride in their diverse student body and faculty.

@@

This book also aims to explain the reasons behind the admission process. There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc. Understanding the reason and mindset of the admission committee and profs. can help you prepare better.

While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
